European or American? A re-examination of the transatlantic influence over the IMF

### 2025-07-25 manuscript/manuscript.qmd
Add European para

Quantitative studies also provide evidence of great power influence over IMF lending decisions. Such studies often use a variety of proxies to measure the recipient country's ties to the major powers, such as UN voting alignment [@thacker1999; @dreher2009], bilateral trade intensity [@barro2005], and financial exposure [@oatley2004]. These proxies aim to capture the strategic incentives of the great powers to support IMF lending to countries that align with its geopolitical interests. Such ties have been proven to be associated with more favourable IMF lending outcomes, such as larger loan amounts, higher participation rates, as well as less and more lenient conditionality [@lipscy2018; @barro2005; @dreher2009].

>>>>

In comparison, the literature on European influence is notably smaller. The most visible manifestation of European power within the Fund is the tradition of European leadership: all Managing Directors of the IMF since its inception have been European nationals [@kahler2001]. Scholars have also pointed to Europe-backed lending decisions, such as those during the Eurozone crisis (notably Greece) [@lipscy2018; @hodson2014], and loans to former colonies in Africa [@stone2004]. However, while these studies offer important insights, they tend to remain case-specific or anecdotal. Crucially, they stop short of offering a systematic comparison between American and European influence, leaving open the question of how these two centres of power operate within the IMF.
